---
title: ""
author: "François Michonneau"
---


```{r setup, echo=FALSE, include=FALSE, messages=FALSE}
library(tidyverse)
library(hrbrthemes)
```


On April 12 and 13th, The Carpentries hosted our first joint Bug BBQ. The main objective of this event was to get the lessons for the Geospatial and Social Sciences Data Carpentry curricula ready for release. We also invited contributions to all Software and Data Carpentry lessons during this time. A Bug BBQ is a global, distributed event, where the community gathers to work collaboratively on our lessons. The combination of having dedicated time to look at the lessons and many fresh sets of eyes to examine their contents makes these events excellent for moving new lessons forward. For older lessons, it is a good opportunity to go through any old issues and pull requests, especially ones that require some extra thinking, or significant changes to the lesson. 

In addition of the Data Carpentry Geospatial and Social Sciences lessons, several other lessons were involved in the Bug BBQ, with Maintainers for those lessons making themselves available during the two day event to faciliate contributions from the community. The DC R Ecology lesson, DC Python Ecology lesson, SWC R Inflammation, SWC R Gapminder (in Spanish), and The Carpentries Instructor Training lesson were all highly active during the Bug BBQ.

While Bug BBQs are distributed events, they are also the opportunity for local communities to get together, and support new contributors locally. Members of the Carpentries community facilitated four local events: Toronto (Canada), Gainesville, FL (USA), Davis, CA (USA), and Buenos Aires (Argentina).


```{r data, echo=FALSE, message=FALSE, warning=FALSE}

issues <- readr::read_csv("data/issues_bug_bbq.csv")

tidy_issues <- tidyr::gather(issues, event, date,
                             -user, -number, -is_pr,
                             -state, -repo, -organization) %>%
    dplyr::mutate(ymd = format.Date(date, "%Y-%m-%d")) %>%
    dplyr::filter(event != "updated_at")

all_issues <- tidy_issues %>% 
    dplyr::count(ymd, event) %>%
    dplyr::mutate(type = "all")

target_issues <- tidy_issues %>%
    dplyr::filter(grepl("geospatial|socialsci", repo)) %>%
    dplyr::mutate(type = case_when(
                      grepl("geospatial", repo) ~ "geospatial",
                      grepl("socialsci", repo) ~ "socialsci",
                      TRUE ~ NA_character_
                  )) %>% 
    dplyr::count(ymd, event, type)

issues <- dplyr::bind_rows(all_issues,
                           target_issues) %>%
    dplyr::mutate(ymd = as.Date(ymd)) %>%
    dplyr::filter(ymd >= "2018-04-01") %>%
    tidyr::complete(ymd, nesting(event, type), fill = list(n = 0L))

filter_bbq_dates <- . %>% 
    dplyr::filter(date >= "2018-04-12" & date <= "2018-04-14")
    
n_issues <- tidy_issues %>%
    filter_bbq_dates %>% 
    dplyr::summarize(n = n_distinct(repo, number)) %>%
    dplyr::pull(n)

n_target_issues <- tidy_issues %>%
    filter_bbq_dates %>%
    dplyr::filter(grepl("geospatial|socialsci", repo)) %>%
    dplyr::summarize(n = n_distinct(repo, number)) %>%
    dplyr::pull(n)

n_contributors <- tidy_issues %>%
    filter_bbq_dates %>%
    dplyr::summarize(n = n_distinct(user))

```

The Bug BBQ was a success in catalyzing a high volume of activity from the community across all of our lesson repositories. Overall, `r n_issues` issues or pull requests were opened and/or closed during the Bug BBQ. This represents a significant increase in activity above our baseline and shows that our community really pulled together to contribute collectively during these two days.


```{r n_issues, out.width="100%", echo=FALSE, message=FALSE, warning=FALSE}

all_issues %>% 
    ggplot(aes(x = as.Date(ymd), y = n, color = event)) +
    annotate("rect", xmin = as.Date("2018-04-11"),
             xmax = as.Date("2018-04-14"), 
             ymin = 0, ymax = Inf,
             alpha = 0.7,
             fill = "gray70") +
    geom_path() +
    scale_x_date(limits = c(as.Date("2018-04-01"), NA)) +
    ylim(c(0, 100)) +
    labs(
        title = "Number of issues and pull requests created and closed",
        subtitle = "Across all Carpentries repositories for April 2018",
        caption = "Shaded area represents the timing of the Bug BBQ",
        x = "Date",
        y = "Number of issues/pull requests"
    ) + 
    scale_color_manual(name = "Issues & \n Pull Requests",
                       labels = c("Closed", "Open"), 
                       values = c("#d45500", "#88aa00")) +
    theme_ipsum_rc()
```

A large proportion of these issues and pull requests (`r n_target_issues` of `r n_issues`) were made to the Geospatial and Social Science repositories, which are in the process of being prepared for their first publication this year. We had `r n_contributors` unique contributors who interacted with issues and pull requests during the Bug BBQ. This is a huge win for our community! 

```{r, out.width="100%", echo=FALSE, message=FALSE, warning=FALSE}

visits <- readr::read_csv("data/visitors.csv")

visits %>%
    dplyr::group_by(timestamp) %>%
    dplyr::summarize(
               n_unique = sum(uniques),
               n_visitors = sum(count)
           ) %>%
    tidyr::gather(visitor_type,  n, -timestamp) %>% 
    ggplot(aes(x = timestamp, y = n)) +
    annotate("rect", xmin = as.POSIXct("2018-04-11"),
             xmax = as.POSIXct("2018-04-14"), 
             ymin = 0, ymax = Inf,
             alpha = 0.7,
             fill = "gray70") +
    geom_line(aes(color = visitor_type)) +
    theme_ipsum_rc() +
    labs(
        title = "Web traffic across our repositories",
        x = "Date",
        y = "Number of vistors (summed across repositories)",
        caption = "Shaded area represents the timing of the Bug BBQ"
    ) +
    scale_color_manual(name = "",
                       labels = c("Unique", "Total"),
                       values = c("#071159", "#ad6a06"))
```


Thank you to all of you who contributed during the Bug BBQ. We hope you had a great experience working on these lessons and interacting
with the broader Carpentries community. If you have any ideas about improving future Bug BBQ events or making it easier for contributors
to contribute to these lessons, please contact François Michonneau (francois@carpentries.org). 

The Maintainers for the Social Science lessons are currently working closely with staff to finalize the Social Science curriculum for publication. These lessons are scheduled for their first publication on May 7th. The Geospatial curriculum will be released and ready to teach at the end of June. 

Thank you for a successful Bug BBQ. Stay tuned for announcements about upcoming Bug BBQs and other lesson development events. 
